---
title: "Lab_14T"
author: "36-600"
date: "Fall 2022"
output:
  html_document:
    toc: no
    toc_float: no
    theme: spacelab
  pdf_document:
    toc: no
---

This week's lab is very much do-it-yourself. On `Canvas`, under `Files > DATA`, there is a file called `creditDefault.csv`. Your boss wants to know if any of the variables included in this file, apart from `Default` itself, are important for predicting whether a person will default rather than paying back their credit-card balance.

From the information above, you should construct an appropriate answer. Note that regardless of whether you think a non-ML model is sufficient, you should always check to see if an ML model gives much better predictive results. Also note that your boss is not necessarily looking for plots here...but you'd always want to do *some* EDA, if for no other reason than to ensure you don't have missing data or outliers!

---
```{r}
suppressMessages(library(dplyr))
suppressWarnings(library(tidyverse))
suppressWarnings(library(gridExtra))
suppressMessages(library(corrplot))
suppressMessages(library(GGally))
suppressMessages(library(FNN))
```
```{r}
df = read.csv("C:/Users/Jie Sing Yoo/Desktop/CMU Coursework/Overview of Statistical Learning and Modeling/R_labs/data/creditDefault.csv",header=TRUE,stringsAsFactors = TRUE)
```

```{r}
head(df)
```

```{r}
summary(df)
```

```{r}
dim(df)
````
```{r}
names(which(colSums(is.na(df))>=1))
```

## EDA

Quantitative Values

```{r}
num_df <- subset(df,select=-c(Student,Default)) %>% gather(.)
```

```{r}
ggplot(data=num_df,mapping=aes(x=value)) +
geom_histogram(color="blue",fill="yellow",bins=25) + 
facet_wrap(~key,scales='free_x') 
```

```{r}
subset(df,select=-c(Student,Default)) %>% cor(.) %>%
corrplot.mixed(.,upper="number",lower="color",tl.pos='lt',diag='l')
```

## Nonquantitative value

```{r}
df_cat = df %>% select(.,Student,Default)  %>% gather(.)
head(df_cat,3)
```


