---
title: "36-600 Project 1"
author: "Jie Sing Yoo"
date: "Fall 2022"
output:
  html_document:
    df_print: paged
---
```{r}
suppressMessages(library(dplyr))
suppressWarnings(library(tidyverse))
suppressWarnings(library(gridExtra))
suppressMessages(library(corrplot))
suppressMessages(library(GGally))
```

## Dataset 

```{r}
df = read.csv("C:/Users/Jie Sing Yoo/Downloads/CMU Coursework/Overview of Statistical Learning and Modeling/R_labs/data/creditBalance.csv",header=TRUE)
```

```{r}
head(df)
```
```{r}
summary(df)
```

No obvious non-informative columns.
No weird values, therefore no error values.

```{r}
names(which(colSums(is.na(df))>=1))
```
This shows that none of the columns has null values.


```{r}
hist(df$Balance,main='Histogram for Credit Balance',xlab='Credit Balance') 
```

## Nonquantitative value

```{r}
df_bar = df %>% select(.,Gender,Married,Student) %>% gather(.)
head(df_bar,3)
```
```{r}
ggplot(df_bar,mapping=aes(x=value)) + geom_bar(color='red',fill='blue') + facet_wrap(~key,scales='free_x')
# People that are non a student are more than people that are student
```

```{r}
plot1 = ggplot(data=df,mapping=aes(x=Gender,y=Balance)) + geom_boxplot(color='black',fill='yellow')
plot2 = ggplot(data=df,mapping=aes(x=Married,y=Balance)) + geom_boxplot(color='black',fill='yellow')
plot3 = ggplot(data=df,mapping=aes(x=Student,y=Balance)) + geom_boxplot(color='black',fill='yellow')
grid.arrange(plot1,plot2,plot3,ncol=3)
```
Credit balance for male and female has no significant difference.
Credit balance for married or not married has no significant difference.
This shows that gender and marital status has no linear correlation with the credit balance.
The credit balance for students is higher on average than those that are not student. However, the sample size for student is too small to draw a confident conclusion.

## Quantitative Value

```{r}
df_new <- subset(df,select=-c(Gender,Student,Married,Balance)) %>% gather(.)
```

```{r}
ggplot(data=df_new,mapping=aes(x=value)) +
geom_histogram(color="blue",fill="yellow",bins=25) + 
facet_wrap(~key,scales='free_x') 
```

```{r}
num.var = 6
ggplot(data=df_new,mapping=aes(x=value,y=rep(df$Balance,num.var))) +
geom_point(size=0.1) +
facet_wrap(~key,scales='free_x') +
ylab("Credit Balance")

```

```{r}
df %>%
subset(.,select=-c(Gender,Student,Married))%>%
cor(.) %>%
corrplot.mixed(.,upper='number',lower="color",tl.pos='lt',diag='l')
```
Limit and Rating of a person in the dataset have high linear correlation with the credit balance.
```{r}
df %>%
subset(.,select=-c(Gender,Student,Married))%>%
ggpairs(.,upper=list(continuous='points'))
```

